Motional Ground-State Cooling Outside the Lamb-Dicke Regime

### 2017-07-09 paper.tex
More numbers corrections and explaination of experimental correction on top of simulation

Our experiment begins by loading a single sodium atom into an optical tweezer from a MOT
for \fxnote{??} ms~\cite{Hutzler2017-LightShifts} and repeats at 2Hz\fxnote{?}.
The tweezer is created by focusing a $700$nm laser beam through an NA=$0.55$ objective to
a beam waist of 0.7 $\mu$m\fxnote{?}.
With $45$mW of tweezer power, the trapping frequencies are
$\{\omega_1,\omega_2,\omega_3\}/2\pi = \{69(1), 430(4), 590(5)\}\ \text{\text{kHz}}$.
Upon loading, an image is taken for $1.5$ms to determine its success.
The same imaging light is used for polarization gradient cooling (PGC) and
reduces the temperature of the single atom to $70$ $\mu$K,
which corresponds to an initial motional state along the three different axes of
$\{\bar n_1, \bar n_2, \bar n_3\}=\{21(2),\, 3.4(2),\, 2.5(2)\}$ (initial $P_0=0.3\%$) in the trap.

@@

Taking the large motional-state changing heating and cooling sources into account,
it was not immediately clear that efficient cooling can be achieved.
We therefore use a Monte-Carlo simulation to guide our search and
to find a robust cooling sequence \fxnote{cite sm}.
In the simulation, we indeed observe a high heating rate due to the large Lamb-Dicke parameters
and confirm that cooling with higher-order Raman sidebands can suppress heating.
In particular, we find that instead of cooling on only one sideband order repeatedly,
it is more efficient to alternate the cooling pulses (Fig.~\ref{f-setup}C) between two
neighboring orders for the axial direction and 2nd- and 1st-orders for the radial directions
to minimize the accumulation of the atom in motional states that have zero Raman coupling.
The simulation also indicates setting the coupling strength of each cooling sideband
to drive a Rabi $\pi$ pulse corresponding to the maximum matrix element motional state
(i.e. the maxima in \ref{f-ld}\fxnote{A or B?}).
Such choice insures each order cools a range of states that have the strongest
couplings and leaves other sidebands to cool the rest.
The efficiency of cooling on higher-order sidebands diminishes
as the atom approaches the ground state.
Therefore, the final cooling cycles utilize only the first-order sideband
while alternating between different axes.

Guided by the promising simulation results,
we construct our axial part of cooling sequence by starting at the highest
two observed cooling sidebands (8th- and 7th-orders)
and decreasing the orders after 6 to 15 cycles.
This sequence gives good initial cooling performance which is further optimized experimentally
by fine tuning of different parameters. \fxnote{which parameters? Elaborate more about the sequence including starting with the second-order radial to remove the most amount of energy initially.}

>>>>

Our experiment begins by loading a single sodium atom into an optical tweezer from a MOT
for $250$ ms~\cite{Hutzler2017-LightShifts} and repeats at $2.5$ Hz.
The tweezer is created by focusing a $700$nm laser beam through an NA=$0.55$ objective to
a beam waist of $0.7$ $\mu$m\fxnote{?}.
With $45$mW of tweezer power, the trapping frequencies are
$\{\omega_1,\omega_2,\omega_3\}/2\pi = \{69(1), 430(4), 590(5)\}\ \text{\text{kHz}}$.
Upon loading, an image is taken for $1.5$ms to determine its success.
The same imaging light is used for polarization gradient cooling (PGC) and
reduces the temperature of the single atom to $70$ $\mu$K,
which corresponds to an initial motional state along the three different axes of
$\{\bar n_1, \bar n_2, \bar n_3\}=\{21(2),\, 3.4(2),\, 2.5(2)\}$ (initial $P_0=0.3\%$) in the trap.

@@

Taking the large motional-state changing heating and cooling sources into account,
it was not immediately clear that efficient cooling can be achieved.
We therefore use a Monte-Carlo simulation to guide our search and
to find a robust cooling sequence\footnote{See supplemental material}.
In the simulation, we indeed observe a high heating rate due to the large Lamb-Dicke parameters
and confirm that cooling with higher-order Raman sidebands can suppress heating.
In particular, we find that instead of cooling on only one sideband order repeatedly,
it is more efficient to alternate the cooling pulses (Fig.~\ref{f-setup}C) between two
neighboring orders for the axial direction and 2nd- and 1st-orders for the radial directions
to minimize the accumulation of the atom in motional states that have zero Raman coupling.
The simulation also indicates setting the coupling strength of each cooling sideband
to drive a Rabi $\pi$ pulse corresponding to the maximum matrix element motional state
(i.e. the maxima in \ref{f-ld}\fxnote{A or B?}).
Such choice insures each order cools a range of states that have the strongest
couplings and leaves other sidebands to cool the rest.
The efficiency of cooling on higher-order sidebands diminishes
as the atom approaches the ground state.
Therefore, the final cooling cycles utilize only the first-order sideband
while alternating between different axes.

Guided by the promising simulation results,
we construct our axial cooling sequence by starting at the highest
two observed cooling sidebands (8th- and 7th-orders)
and decreasing the orders after $6$ to $15$ cycles.
This process is repeated until we reach the first order cooling sideband where most of the
population are in the first few excited state at which point we switch to only cooling on the
first order sideband with two different pulse lengths in order to efficiently cool atoms in the
few remaining motional states.
The radial cooling is done similarly to the axial cooling with the initial cooling orders being
the 2nd and 1st order and switching to first order only after $20$ to $30$ pulses.
This sequence gives good initial cooling performance which is used to measure experimental
parameters, including the Rabi rates of the Raman beams and optical pumping rates,
in order to optimize the sequence further.
